---
title: "Changes in gene DNA methylation and expression networks accompany caste specialization and age-related physiological changes in a social insect"
author: "Morandin et al."
date: "21/02/2019"
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(WGCNA)
```

# Module preservation
## Preservation of the co-expression network within the co-methylation network
### Input both datasets
```{r preservation1}
dat0=read.table("Preservation_Expression.txt", header=T)
head(dat0)
datSummary=dat0[,c(1:2)]
datExpr <- t(dat0[3:18])
no.samples <- dim(datExpr)[[1]]
dim(datExpr)
colnames(datExpr) = datSummary$cluster 
colorsExp = dat0$color

dat=read.table("Preservation_methylation.txt", header=T)
head(dat)
datMetSummary=dat[,c(1:2)]
datMet <- t(dat[3:22])
no.samples <- dim(datMet)[[1]]
dim(datMet)
colnames(datMet) = datMetSummary$cluster 
colorsMet = dat$color
```

### Calculation Module preservation
```{r calculation1}
setLabels = c("Met", "Exp");
multiExpr = list(Exp = list(data = datExpr), Met = list(data = datMet));
multiColor = list(Exp = colorsExp);

system.time( {
  mp = modulePreservation(multiExpr, multiColor,
                          referenceNetworks = 1,
                          nPermutations = 100,
                          calculateQvalue = TRUE,
                          randomSeed = 1,
                          quickCor = 0,
                          verbose = 3)
} );
```

### Save outputs
```{r save1}
save(mp, file = "modulePreservation_metExp.RData");
```

#### Analysis and Display
```{r display1}
ref = 1
test = 2
statsObs = cbind(mp$quality$observed[[ref]][[test]][, -1], mp$preservation$observed[[ref]][[test]][, -1])
statsZ = cbind(mp$quality$Z[[ref]][[test]][, -1], mp$preservation$Z[[ref]][[test]][, -1]);

# Compare preservation to quality:
print( cbind(statsObs[, c("medianRank.pres", "medianRank.qual")],
             signif(statsZ[, c("Zsummary.pres", "Zsummary.qual")], 2)) )

# Module labels and module sizes are also contained in the results
modColors = rownames(mp$preservation$observed[[ref]][[test]])
moduleSizes = mp$preservation$Z[[ref]][[test]][, 1];

# leave grey and gold modules out
plotMods = !(modColors %in% c("gold","grey"));

# Text labels for points
text = modColors[plotMods];

# Auxiliary convenience variable
plotData = cbind(mp$preservation$observed[[ref]][[test]][, 2], mp$preservation$Z[[ref]][[test]][, 2])

# Plot
mains = c("Preservation Median rank", "Preservation Zsummary");
sizeGrWindow(10, 5);
par(mfrow = c(1,2))
par(mar = c(4.5,4.5,2.5,1))
for (p in 1:2)
{
  min = min(plotData[, p], na.rm = TRUE);
  max = max(plotData[, p], na.rm = TRUE);
  # Adjust ploting ranges appropriately
  if (p==2)
  {
    if (min > -max/10) min = -max/10
    ylim = c(min - 0.1 * (max-min), max + 0.1 * (max-min))
  } else
    ylim = c(max + 0.1 * (max-min), min - 0.1 * (max-min))
  plot(moduleSizes[plotMods], plotData[plotMods, p], col = 1, bg = modColors[plotMods], pch = 21,
       main = mains[p],
       cex = 1.4,
       ylab = mains[p], xlab = "Module size", log = "x",
       ylim = ylim,
       xlim = c(1, 4000), cex.lab = 1.2, cex.axis = 1.2, cex.main =1.4)
  if (p==2) {
    abline(h=0)
    abline(h=2, col = "blue", lty = 2)
    abline(h=10, col = "darkgreen", lty = 2)
  } }
```



## Preservation of the co-methylation network within the co-expression network
### Calculation Module preservation
```{r calculation2}
setLabels = c("Met", "Exp");

multiExpr = list(Met = list(data = datMet),Exp = list(data = datExpr));
multiColor = list(Met = colorsMet);

system.time( {
  mp = modulePreservation(multiExpr, multiColor,
                          referenceNetworks = 1,
                          nPermutations = 100,
                          randomSeed = 1,
                          quickCor = 0,
                          verbose = 3)
} );
```

### Save outputs
```{r save2}
save(mp, file = "modulePreservation_metExp_reverse.RData");
```

#### Analysis and Display
```{r display2}
ref = 1
test = 2
statsObs = cbind(mp$quality$observed[[ref]][[test]][, -1], mp$preservation$observed[[ref]][[test]][, -1])
statsZ = cbind(mp$quality$Z[[ref]][[test]][, -1], mp$preservation$Z[[ref]][[test]][, -1]);

# Compare preservation to quality:
print( cbind(statsObs[, c("medianRank.pres", "medianRank.qual")],
             signif(statsZ[, c("Zsummary.pres", "Zsummary.qual")], 2)) )

# Module labels and module sizes are also contained in the results
modColors = rownames(mp$preservation$observed[[ref]][[test]])
moduleSizes = mp$preservation$Z[[ref]][[test]][, 1];

# leave grey and gold modules out
plotMods = !(modColors %in% c("gold","grey"));

# Text labels for points
text = modColors[plotMods];

# Auxiliary convenience variable
plotData = cbind(mp$preservation$observed[[ref]][[test]][, 2], mp$preservation$Z[[ref]][[test]][, 2])

# Plot
mains = c("Preservation Median rank", "Preservation Zsummary");
sizeGrWindow(10, 5);
par(mfrow = c(1,2))
par(mar = c(4.5,4.5,2.5,1))
for (p in 1:2)
{
  min = min(plotData[, p], na.rm = TRUE);
  max = max(plotData[, p], na.rm = TRUE);
  # Adjust ploting ranges appropriately
  if (p==2)
  {
    if (min > -max/10) min = -max/10
    ylim = c(min - 0.1 * (max-min), max + 0.1 * (max-min))
  } else
    ylim = c(max + 0.1 * (max-min), min - 0.1 * (max-min))
  plot(moduleSizes[plotMods], plotData[plotMods, p], col = 1, bg = modColors[plotMods], pch = 21,
       main = mains[p],
       cex = 1.4,
       ylab = mains[p], xlab = "Module size", log = "x",
       ylim = ylim,
       xlim = c(1, 4000), cex.lab = 1.2, cex.axis = 1.2, cex.main =1.4)
  if (p==2) {
    abline(h=0)
    abline(h=2, col = "blue", lty = 2)
    abline(h=10, col = "darkgreen", lty = 2)
  } }
```

## Preservation of the co-expression network within the Morandin et al. (2016) network
### Input both datasets
```{r input2}
dat0=read.table("Preservation_Morandin2016.txt", header=T)
head(dat0)
datSummary=dat0[,c(1:2)]
datOld <- t(dat0[3:102])
no.samples <- dim(datOld)[[1]]
dim(datOld)
colnames(datOld) = datSummary$cluster 
colorsOld = dat0$color 

dat=read.table("Preservation_Expression.txt", header=T)
head(dat)
datExpSummary=dat[,c(1:2)]
datExp <- t(dat[3:18])
no.samples <- dim(datExp)[[1]]
dim(datExp)
colnames(datExp) = datExpSummary$cluster 
colorsExp = dat$color 
```

### Calculation Module preservation
```{r calculation3}
setLabels = c("Old", "Exp");
multiExpr = list(Old = list(data = datOld), Exp = list(data = datExp));
multiColor = list(Old = colorsOld);

system.time( {
  mp = modulePreservation(multiExpr, multiColor,
                          referenceNetworks = 1,
                          nPermutations = 100,
                          calculateQvalue = TRUE,
                          randomSeed = 1,
                          quickCor = 0,
                          verbose = 3)
} );
```

### Save outputs
```{r save3}
save(mp, file = "modulePreservation_OldExp.RData");
```

#### Analysis and Display
```{r display3}
ref = 1
test = 2
statsObs = cbind(mp$quality$observed[[ref]][[test]][, -1], mp$preservation$observed[[ref]][[test]][, -1])
statsZ = cbind(mp$quality$Z[[ref]][[test]][, -1], mp$preservation$Z[[ref]][[test]][, -1]);

# Compare preservation to quality:
print( cbind(statsObs[, c("medianRank.pres", "medianRank.qual")],
             signif(statsZ[, c("Zsummary.pres", "Zsummary.qual")], 2)) )

# Module labels and module sizes are also contained in the results
modColors = rownames(mp$preservation$observed[[ref]][[test]])
moduleSizes = mp$preservation$Z[[ref]][[test]][, 1];

# leave grey and gold modules out
plotMods = !(modColors %in% c("gold","grey"));

# Text labels for points
text = modColors[plotMods];

# Auxiliary convenience variable
plotData = cbind(mp$preservation$observed[[ref]][[test]][, 2], mp$preservation$Z[[ref]][[test]][, 2])

# Plot
mains = c("Preservation Median rank", "Preservation Zsummary");
sizeGrWindow(10, 5);
par(mfrow = c(1,2))
par(mar = c(4.5,4.5,2.5,1))
for (p in 1:2)
{
  min = min(plotData[, p], na.rm = TRUE);
  max = max(plotData[, p], na.rm = TRUE);
  # Adjust ploting ranges appropriately
  if (p==2)
  {
    if (min > -max/10) min = -max/10
    ylim = c(min - 0.1 * (max-min), max + 0.1 * (max-min))
  } else
    ylim = c(max + 0.1 * (max-min), min - 0.1 * (max-min))
  plot(moduleSizes[plotMods], plotData[plotMods, p], col = 1, bg = modColors[plotMods], pch = 21,
       main = mains[p],
       cex = 1.4,
       ylab = mains[p], xlab = "Module size", log = "x",
       ylim = ylim,
       xlim = c(1, 4000), cex.lab = 1.2, cex.axis = 1.2, cex.main =1.4)
  if (p==2) {
    abline(h=0)
    abline(h=2, col = "blue", lty = 2)
    abline(h=10, col = "darkgreen", lty = 2)
  } }
```