---
title: "Changes in gene DNA methylation and expression networks accompany caste specialization and age-related physiological changes in a social insect"
author: "Morandin et al."
date: "19/02/2019"
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(flashClust)
library(WGCNA)
library(dynamicTreeCut)
library(psych)
library(WGCNA)
library(Hmisc)
library(tidyverse)
library(boot)
library(boot)
library(magrittr)
options(stringsAsFactors = FALSE);
```

# Co-methylation network analysis
## Loading and cleaning the data
```{r cleaning}
dat1=read.table("Comethylation_input.txt", header=T)
dim(dat1)
names(dat1)
datExpr0 = as.data.frame(t(dat1[, -c(1)]));
names(datExpr0) = dat1$Genes;
rownames(datExpr0) = names(dat1)[-c(1)];
```

### Quality filtering step
```{r filtering}
# Remove genes with too many missing values, to be retained a gene needs to have values in at least half of the samples.
gsg = goodSamplesGenes(datExpr0, verbose = 3)
gsg$allOK

if (!gsg$allOK)
{
  if (sum(!gsg$goodGenes)>0)
    printFlush(paste("Removing genes:", paste(names(datExpr0)[!gsg$goodGenes], collapse = ", ")));
  if (sum(!gsg$goodSamples)>0)
    printFlush(paste("Removing samples:", paste(rownames(datExpr0)[!gsg$goodSamples], collapse = ", ")));
  datExpr0 = datExpr0[gsg$goodSamples, gsg$goodGenes]
}
```

### Cluster the samples
```{r cluster}
# Cluster the samples to inspect for outlier arrays
sampleTree = flashClust(dist(datExpr0), method = "average");
plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 1.5,cex.axis = 1.5, cex.main = 2)
abline(h = 2000, col = "red");
clust = cutreeStatic(sampleTree, cutHeight = 4000000, minSize = 10)
table(clust)
keepSamples = (clust==1)
datExpr = datExpr0[keepSamples, ]
nGenes = ncol(datExpr)
nSamples = nrow(datExpr)
```

### Load trait data
```{r trait}
traitData = read.table("Comethylation_trait.txt", header=T);
dim(traitData)
names(traitData)
allTraits = traitData;
dim(allTraits)
names(allTraits)

# Form a data frame that will hold the traits
WWDSDRSamples = rownames(datExpr);
traitRows = match(WWDSDRSamples, allTraits$Samples);
datTraits = allTraits[traitRows, -1];
rownames(datTraits) = allTraits[traitRows, 1];
collectGarbage();

#Re-cluster the samples
sampleTree2 = flashClust(dist(datExpr), method = "average")
traitColors = numbers2colors(datTraits, signed = FALSE);
plotDendroAndColors(sampleTree2, traitColors, groupLabels = names(datTraits), main = "Sample dendrogram and trait heatmap")
save(datExpr, datTraits, file = "Comethylation_part1.RData")

```

## Network Construction and Module Detection
### Choose the soft-thresholding power for analysis of network topology
```{r Constrution}
lnames = load(file = "Comethylation_part1.RData");
lnames
powers = c(c(1:10), seq(from = 12, to=40, by=1))

# Call the network topology analysis function
sft = pickSoftThreshold(datExpr, powerVector = powers, verbose = 5)
sizeGrWindow(9, 5)
cex1 = 1;
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit, signed R^2", type="n", main = paste("Scale independence"));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], labels=powers,cex=cex1,col="red");
abline(h=0.7,col="red")
plot(sft$fitIndices[,1], sft$fitIndices[,5], xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n", main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex.axis = 3,col="green",cex.lab =20,cex.main =2)
```

### Calculate the adjacencies using the soft thresholding power
```{r adjancencies}
softPower = 20;
adjacency = adjacency(datExpr, power = softPower);
TOM = TOMsimilarity(adjacency); 
dissTOM = 1-TOM
geneTree = flashClust(as.dist(dissTOM), method = "average");
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity", labels = FALSE, hang = 0.04);
```

### Module identification using dynamic tree cut
```{r tree cut}
minModuleSize = 30;
dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM, deepSplit = 2, cutHeight = 0.97, pamRespectsDendro = FALSE, minClusterSize = minModuleSize);
table(dynamicMods)
dynamicColors = labels2colors(dynamicMods)
table(dynamicColors)
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut", dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05, main = "Gene dendrogram and module colors")
```

### Merge modules with similar methylation profiles 
```{r merge}
MEList = moduleEigengenes(datExpr, colors = dynamicColors,excludeGrey = TRUE)
MEs = MEList$eigengenes
MEDiss = 1-cor(MEs);
METree = flashClust(as.dist(MEDiss), method = "average");
plot(METree, main = "Clustering of module eigengenes", xlab = "", sub = "")
MEDissThres = 0.45
abline(h=MEDissThres, col = "red")
merge = mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres, verbose = 3)
mergedColors = merge$colors;
table(mergedColors)
mergedMEs = merge$newMEs;
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors), c("Dynamic Tree Cut", "Merged dynamic"),
                    dendroLabels = FALSE, hang = 0.03, addGuide = TRUE, guideHang = 0.05)
moduleColors = mergedColors
colorOrder = c(standardColors(50));
moduleLabels = match(moduleColors, colorOrder)-1;
MEs = mergedMEs;

save(MEs, moduleLabels, moduleColors, geneTree, file = "Comethylation_part2.RData")
```

### Bootstrap glm to correlate eigengenes and traits
```{r glm}
## Input Data ##
dat<-na.omit(read_tsv("Comethylation_glm.txt"))
dat.tidy <- dat %>% gather(eigengene, eigenvalue, -ID, -Caste, -Stage) 

# Bootstrap glm #
bs <- function(formula, data, indices) {
  d <- data[indices,] # allows boot to select sample 
  fit <- lm(formula, data=d)
  return(coef(fit,use = "complete.obs"))
}

## M1 ##
results <-  boot(data=subset(dat.tidy, eigengene == "MEsienna3"), statistic=bs, R=1000, formula=eigenvalue ~ Caste*Stage)

# Confidence intervals
boot.ci(results, type="bca", index=2) # Caste
boot.ci(results, type="bca", index=3) # Stage
boot.ci(results, type="bca", index=4) # Both

# Extrapolate pvalue
## Caste
conf <- c()
alphas <- seq(1,.001,by=-0.001)
for (i in alphas) {
  conf <- c(conf, boot.ci(results, type="bca", index=2, conf=1-i)$bca[5])
}
ggplot(data.frame(alphas=alphas,conf=conf), aes(conf,alphas))+geom_line() +geom_vline(xintercept=0)
data.frame(alphas=alphas,conf=conf); df[length(df$conf < 0)+1,]

## Stage
conf <- c()
alphas <- seq(1,.001,by=-0.001)
for (i in alphas) {
  conf <- c(conf, boot.ci(results, type="bca", index=3, conf=1-i)$bca[5])
}
ggplot(data.frame(alphas=alphas,conf=conf), aes(conf,alphas))+geom_line() +geom_vline(xintercept=0)
data.frame(alphas=alphas,conf=conf); df[length(df$conf < 0)+1,]

## Both
conf <- c()
alphas <- seq(1,.001,by=-0.001)
for (i in alphas) {
  conf <- c(conf, boot.ci(results, type="bca", index=4, conf=1-i)$bca[4])
}
ggplot(data.frame(alphas=alphas,conf=conf), aes(conf,alphas))+geom_line() +geom_vline(xintercept=0)
data.frame(alphas=alphas,conf=conf); df[length(df$conf < 0)+1,]

## Can be done for all modules
### pvalues were FDR corrected afterwards
```
